Leader functions and qualities across 59 cultures: an exploratory analysis

### 2019-11-30 leadership_across_cultures_contexts.Rmd
Minor edits to exploratory model plots.
Working on MS.

Added visreg exploratory model effects plots to MS to visualize effects. Might remove these plots/analyses prior to submission.

Leadership and followership are critical in understanding group structure and have been the focus of much scientific attention across psychology, sociology, political science, and managerial studies. Until recently, this very large body of work was relatively disconnected from approaches to leadership in anthropology and biology. Under a unifying evolutionary framework, social and biological scientists have now produced a sizable body of theory on the evolution and cross-cultural patterning of leadership and followership [e.g., @henrich_big_2015; @richerson_tribal_2009; @van_vugt_evolutionary_2006; @pietraszewski_evolution_2019; @hooper_theory_2010; @kaplan_evolutionary_2009; @von_rueden_roots_2014], accompanied by a growing body of empirical results [e.g., @bowser_womens_2010; @cheng_two_2013; @garfield_investigating_2019; @garfield_evolutionary_2019; @von_rueden_leadership_2014; @von_rueden_sex_2018; @smith_leadership_2016]. This integration as been informative. Drawing on the study of leadership in evolutionary perspective and across diverse, small-scale and traditional societies can help develop more generalizable theories and offer applications to shape organizational structure in ways which are more consistent with human evolutionary history and evolved leader-follower psychologies [@von_rueden_leadership_2015; @garfield_review_2019]. Nonetheless, much of this work has focused on community or political leaders, often overlooking the nature of leadership and followership across social domains and at all levels of social organization. Moreover, while the benefits leaders accrue for their services have received substantial attention the costs of leadership for both leaders and followers have also been relatively under-investigated. Current outstanding questions for evolutionary leadership scholars include the degree to which the nature of leadership is shaped by population-level variation and the costs and benefits leaders and followers incur across distinct leadership systems.

@@

```{r, fig.height=20, fig.width=9}
# qc_bayes_model_post_plot
```

@@

## The cultural dependancy of human leadership

The distribiution of text records by group type and subsistence suggests leaders emerge within similar social domains among populations within varied subsistance strategies, but at variable frequencies. Hunter-gatherers and pastoralists, populations more likely to expeirence high residential mobility exhibited greater relative frequencies of leaderhsip wtihin residential groups and kin groups. Pastoralists, however, rely in part on a highly defensible resource base (i.e., livestock) and exhibit leadership beyond the community mores o than hunter-gatherers and on par with horticulturalists. Both pastoralists and horticulturalists are known for elevated rates of inter-group conflict relative to mobile hunter-gatherers and these results are consistent with a relationship between increased inter-group conflict and more elaboratored political leadership systems.

@@

>>>>

Leadership and followership are critical in understanding group structure and have been the focus of much scientific attention across psychology, sociology, political science, and managerial studies. Until recently, this very large body of work was relatively disconnected from approaches to leadership in anthropology and biology. Under a unifying evolutionary framework, social and biological scientists have now produced a sizable body of theory on the evolution and cross-cultural patterning of leadership and followership [e.g., @price_evolution_2014; @henrich_big_2015; @richerson_tribal_2009; @van_vugt_evolutionary_2006; @pietraszewski_evolution_2019; @hooper_theory_2010; @kaplan_evolutionary_2009; @von_rueden_roots_2014; @hagen_leadership_2019], accompanied by a growing body of empirical results [e.g., @bowser_womens_2010; @cheng_two_2013; @garfield_investigating_2019; @garfield_evolutionary_2019; @von_rueden_leadership_2014; @von_rueden_sex_2018; @smith_leadership_2016; @smith_obstacles_2018]. This integration as been informative. Drawing on the study of leadership in evolutionary perspective and across diverse, small-scale and traditional societies can help develop more generalizable theories and offer applications to shape organizational structure in ways which are more consistent with human evolutionary history and evolved leader-follower psychologies [@von_rueden_leadership_2015; @garfield_review_2019]. Nonetheless, much of this work has focused on community or political leaders, often overlooking the nature of leadership and followership across social domains and at all levels of social organization. Moreover, while the benefits leaders accrue for their services have received substantial attention the costs of leadership for both leaders and followers have also been relatively under-investigated. Current outstanding questions for evolutionary leadership scholars include the degree to which the nature of leadership is shaped by population-level variation and the costs and benefits leaders and followers incur across distinct leadership systems.

@@

Functions PC 1 model

```{r funcPCmodel1, fig.height=8, fig.width=20}
m_fPC1_sub_plot + m_fPC1_group_plot
```

Functions PC 2 model

```{r funcPCmodel2, fig.height=10, fig.width=20}
m_fPC2_sub_plot + m_fPC2_group_plot
```

Qualities PC 1 model

```{r qualPCmodel1, fig.height=10, fig.width=20}
m_qPC1_sub_plot + m_qPC1_group_plot
```

Qualities PC 2 model

```{r qualPCmodel2, fig.height=10, fig.width=20}
m_qPC2_sub_plot + m_qPC2_group_plot
```

@@

## The cultural and context dependancy of human leadership

The distribiution of text records by group type and subsistence suggests leaders emerge within similar social domains among populations within varied subsistance strategies, but at variable frequencies. Hunter-gatherers and pastoralists, populations more likely to expeirence high residential mobility, exhibited greater relative frequencies of leaderhsip wtihin residential groups and kin groups. Pastoralists, however, rely in part on a highly defensible resource base (i.e., livestock) and exhibit leadership beyond the community more so than hunter-gatherers and on par with horticulturalists. Pastoral and horticultural populations are often experience elevated rates of inter-group conflict relative to mobile hunter-gatherers and these results are consistent with a relationship between increased inter-group conflict and more elaboratored political leadership systems.

@@

# Conclusion
